### 2021-09-17 products/manuscript/Manuscript.Rmd
Completed Part 1 for Project
Completed the first part for the MADA project. The raw data files are in the raw data folder. The main text and glimpses of the data in code are in the "Manuscript.Rmd" file in the products folder.

---
title: "MADA Project - Ibis Blood Parasites from Florida"
subtitle: ""
author: Ryan Grunert
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: article
output:
  bookdown::word_document2: 
    toc: false
    number_sections: true
  bookdown::html_document2: 
    toc: false
bibliography: ../references.bib
csl: ../apa.csl
---
```{r}
#Loading required packages
library(dplyr) #for data processing
library(here) #to set paths
library(tidyverse) #all required data manipulation packages
library(xlsx) #to read in excel spreadsheets
```

# Introduction

## General Background Information
The data I will be using for this project is American White Ibis (*Eudocimus albus*) blood parasite data that was collected from 2010-2017. The parasite the data focuses on is the protozoan *Haemoproteus*, which causes a chronic blood infection in White Ibises and many other birds. *Haemoproteus* is a vector-borne disease, it is transmitted by *Culicoides* biting midges when they feed on the birds and causes long-lasting parasitemia. Adult healthy birds tolerate haemosporidian infections well, but young birds and birds outside their normal range that are naive to the pathogen are more likely to develop clinical disease. Haemosporidian infections usually cause long-lasting chronic infections in natural hosts, and can cause reduced reproductive success, host fitness, and increased stress and disease susceptibility.

Here is a link to a previously published study concerning this data -> https://www.sciencedirect.com/science/article/pii/S221322441730072X

## Description of data and data source
The data was collected in a collaboration between Dr. Michael Yabsley's lab (of which I am a part) and Dr. Sonia Hernandez's lab here at UGA. I received the raw data from Dr. Yabsley in an agreement to analyze the data over the course of the semester in exchange for using the data for the MADA project.

The data is organized in excel spreadsheets. The two main spreadsheets (Ibis_Blood_ParasiteData_2017 and IbisFieldDataOct19_2017) hold the majority of the data, and two supplemental spreadsheets from other collaborators that were or currently working on the project contain additional variables that will be combined with the main dataset in R for additional analyses. Each observation in the data is an individual White Ibis that was captured and sampled, and each variable is a different descriptor of either the Ibis itself, or the site/environment it was captured in.

Two datasheets have overlapping information (Ibis_Blood_ParasiteData_2017 and IbisFieldDataOct19_2017), but both will be combined to have a complete dataset for analysis. The first contains parasitemia data, while the second contains field data. The 100IbisID_UrbanGradient datasheet will be used to assist with the analysis concerning *Haemoproteus* presence and habitat type. The SalmonellaFieldSerotypes datasheet may be used to compare the presence of both *Haemoproteus* and *Salmonella* in the Ibis, although that may not be needed considering the difference in biology and epidemiology of the two pathogens.

```{r}
#Setting the path to the Ibis blood parasite spreadsheet
data_locationIbisB <- here::here("data","raw_data","Ibis_Blood_ParasiteData_2017.xlsx")

#Loads the raw Ibis blood data excel sheet from 2015-2017
rdIbisBlood15_17 <- read.xlsx(data_locationIbisB, sheetName = "2015-2017")

#Loads the raw Ibis blood data excel sheet from 2010-2014
rdIbisBlood10_14 <- read.xlsx(data_locationIbisB, sheetName = "2010-2014")

glimpse(rdIbisBlood15_17)
glimpse(rdIbisBlood10_14)

```

The main variables I'm interested in for these two sheets above are "Haemoproteus.parasitemia.", and the first "X" column is the sample/bird ID. The 2010-14 set is missing more data, but characteristics of the birds from that group may be used.

```{r}
#Setting the path to the Ibis field data spreadsheet
data_locationIbisF <- here::here("data","raw_data","IbisFieldDataOct19_2017.xlsx")

#Loads the raw Ibis field data excel sheet
rdIbisField <- read.xlsx(data_locationIbisF, sheetName = "All capture data")

glimpse(rdIbisField)

```

This dataset is the most organized sheet, but columns of interest will be extracted from this sheet as well as the others to create new clean data frames for analysis.

```{r}
#Setting the path to the first Ibis supplemental excel sheet (Urban Gradients)
data_locationIbisUr <- here::here("data","raw_data","100IbisID_UrbanGradient.xlsx")

#Loads the raw Ibis Urban Gradient supplemental data sheet
UrbanIbis <- read.xlsx(data_locationIbisUr, sheetName = "Samples by Date")

glimpse(UrbanIbis)

```

These samples from this datasheet will be used in conjunction with the two main datasheets in order to create an "urbanization gradient" for the samples that are located in either rural or urban areas. The gradient will most likely range from least ("1", rural) to most ("4", urban) urban for example, but further work and planning needs to be done first in order to figure out the specifics of the analysis.

```{r}
#Setting the path to the second Ibis supplemental excel sheet (Salmonella)
data_locationIbisSal <- here::here("data","raw_data","SalmonellaFieldSerotypesAug27_2018.xlsx")

#Loads the raw Ibis Urban Gradient supplemental data sheet
SalIbis <- read.xlsx(data_locationIbisSal, sheetName = "ALL Positives")

glimpse(SalIbis)
```

This sheet details the Ibises that tested positive for Salmonella, but this data may not be used because of the differences in pathogenesis and epidemiology of the two diseases.

## Questions/Hypotheses to be addressed
The questions I plan to address are whether the amounts of *Haemoproteus* parasitemia in the various Ibises are significantly different when focusing on age, sex, season, location, or habitat type (Urban and Rural). Does the amount of parasitemia in the birds decrease, increase or remain the same over time as they get older? Is parasitemia significantly higher in birds sampled in urban or rural environments? These are the main variables I will be looking at for this project, but I may analyze others as the project continues. Not all of the variables were collected for every Ibis sampled, and the data still needs cleaning to accurately detail the sample size for each analysis.

## Expected Analyses
I think that I will have to conduct multiple ANOVA's during the analysis process. Linear regression models may apply here too for some variables, but I'm not entirely sure. I would like to explore the possibility of using the location data for each sample to conduct a geographical analysis with a map as the output. I will be continuing to contact those involved with the study as well to see what analyses they want to be conducted and what questions they want answered.

# Methods and Results

## Data aquisition

## Data import and cleaning

## Exploratory analysis

## Full analysis

# Discussion

## Summary and Interpretation

## Strengths and Limitations

## Conclusions

# References

# Summary/Abstract
_Write a summary of your project._